Ultra Diffuse Galaxies associated with Milky-Way Analogs

### 2022-08-25 main.tex
some discussion on UDG formation

@@

\subsection{Abundance of UDGs and UPGs}\label{sec:n_udg}

\begin{figure*}
	\vbox{ 
		\centering
		\includegraphics[width=1\linewidth]{N_UDG_host_mass.pdf}
	}
    \caption{The abundance of UDGs as a function of host halo mass. \textit{Left}: We compile the UDG abundance measurements from literature covering a wide range of host mass. The UDG abundance of our work $N_{\rm UDG} = 0.76\pm 0.04$ is shown as red square, the power-law relation from \citet{vdBurg2017} is shown in gray, and our fitting result is shown in pink. Our UDG abundance is consistent with ELVES but significantly higher than that in SAGA. After including more measurements at the lower halo mass end, the power-law is shallower than the relation reported in \citet{vdBurg2017}. \textit{Right}: We split the UDG sample into bins based on the host halo mass (triangles) and $g-i$ color (pentagons). The UDG abundance is higher for more massive host and redder host. }
    \label{fig:n_udg}
\end{figure*}

@@

We refer to quenching as the cessation of star formation in galaxies. In principle, direct estimates for the instantaneous star formation rate (SFR) from UV or H$\alpha$ are needed to define a quenched galaxy  population. As an example, the SAGA survey measures the H$\alpha$ equivalent width (EW) from spectra to define quenched satellites. However, these specific data sets are either not deep enough to probe our UDGs and UPGs, or they are very expensive to obtain and cover all the footprint of our sample. Therefore, we seek other indicators including integrated color and morphology to approximate star-forming properties. \citet{CarlstenELVES2022} visually inspected all satellites and classified them into early-type (red and smooth) and late-type (blue, asymmetric, clumpy). Using this sample, they find that a mass-dependent color cut $g-i = -0.067 M_V - 0.23$ could also divide the sample into early- and late-type that are nearly identical to the ones based on morphology. The derived quenched fractions based on morphology and color are very similar. Furthermore, \citet{Font2022} showed in simulations that this color cut effectively separates star forming galaxies from quenched ones. Since we do not visually inspect all UDGs and UPGs in our sample, we take the color cut from \citet{CarlstenELVES2022} to define our quenched galaxies: galaxies that are redder than $g-i > -0.067 M_V - 0.23$ are defined as quenched, whereas those bluer galaxies are thought to be star-forming. However, one caveat is that this criterion was derived from LV satellites in \citet{CarlstenELVES2022}, and the bulk of them are brighter than our UDGs and UPGs. It is not guaranteed that the same color cut would work in our regime. \footnote{\jiaxuan{Add some robustness tests, such as ``Changing the intercept of the color cut to -0.28 makes the overall quenched fraction higher, but does not change the trend of quenched fraction as a function of stellar mass.''}}

@@

@@

\subsection{High and constant quenched fraction of UDGs is an artifact}

In Section \ref{sec:quench}, we calculate the quenched fraction as a function of stellar mass for UDGs (left panels in Figure \ref{fig:qfrac}). We find that the UDG quenched fraction is reasonably a constant over $>1$ dex in stellar mass, and is higher than the quenched fractions of UPGs and satellite population of MW analogs. In this subsection, we argue that UDG's high and constant quenched fraction is merely an artifact due to its contrived definition.

@@

\subsection{UDG formation}
Discuss how does quenching and pufffing interplay.

Galaxy conformity: mass-size independent of color, jenny's quenched fraction independent of morph.

Pre-puffed up or pre-quenched?

Amorisco with my size distribution?

@@

Scenarios:
1. violent environmental effects: radial orbits, back-splash both explain large size and quenching. But according to Bernavidis et al, the quenched fraction drops with increasing mass.

2. accreting field UDGs, and quench them due to ram pressure stripping. But how long do accreted UDGs last before being destroyed? Need to look at ram pressure stripping literature.

Homework:
3. Move the stellar mass bin definition, try to match with red/blue, spiral/elliptical plot. If cannot match, does this hint the formation scenarios?
4. How significant do we detect blue UDGs? 
5. Split into redshift bins

UDG rejuvenate after ram pressure stripping?

Why red is equivalent to quenched? Need to clarify in the paper

calculate probability of being a backsplash satellite (e.g., splashed to 1.5 Mpc)

Compare spatial distribution with normal dwarf? iF they are indeed high-spin tail of normal dwarfs.

Don't forget to discuss \citep{xSAGA-I}

@@

\section{Summary}

>>>>

\jiaxuan{We don't discriminate our hosts to be paired or not. 
Don't forget to discuss \citep{xSAGA-I} }

@@

\subsection{UDG and UPG on mass-size plane}

\subsection{Abundance of UDGs and UPGs}\label{sec:n_udg}
\jiaxuan{Discuss NGVS}
\begin{figure*}
	\vbox{ 
		\centering
		\includegraphics[width=1\linewidth]{N_UDG_host_mass.pdf}
	}
    \caption{The abundance of UDGs as a function of host halo mass. \textit{Left}: We compile the UDG abundance measurements from literature covering a wide range of host mass. The UDG abundance of our work $N_{\rm UDG} = 0.76\pm 0.04$ is shown as red square, the power-law relation from \citet{vdBurg2017} is shown in gray, and our fitting result is shown in pink. Our UDG abundance is consistent with ELVES but significantly higher than that in SAGA. After including more measurements at the lower halo mass end, the power-law is shallower than the relation reported in \citet{vdBurg2017}. \textit{Right}: We split the UDG sample into bins based on the host halo mass (triangles) and $g-i$ color (pentagons). The UDG abundance is higher for more massive host and redder host. }
    \label{fig:n_udg}
\end{figure*}

@@

We don't discriminate our hosts to be paired or not.

We refer to quenching as the cessation of star formation in galaxies. In principle, direct estimates for the instantaneous star formation rate (SFR) from UV or H$\alpha$ or estimations for the \ion{H}{1} gas are needed to define a quenched galaxy  population. As an example, the SAGA survey measures the H$\alpha$ equivalent width (EW) from spectra to define quenched satellites. There are also many studies probing \ion{H}{1} masses for MW and M31 satellites \citep[e.g.,][]{Grcevich2009,Spekkens2014,Putman2021}. However, these specific data sets are either not deep enough to probe our UDGs and UPGs, or they are very expensive to obtain and cover all the footprint of our sample. Therefore, we seek other indicators including integrated color and morphology to approximate star-forming properties. \citet{CarlstenELVES2022} visually inspected all satellites and classified them into early-type (red and smooth) and late-type (blue, asymmetric, clumpy). Using this sample, they find that a mass-dependent color cut $g-i = -0.067 M_V - 0.23$ could also divide the sample into early- and late-type that are nearly identical to the ones based on morphology. The derived quenched fractions based on morphology and color are very similar. Furthermore, \citet{Font2022} showed in simulations that this color cut effectively separates star forming galaxies from quenched ones. This makes the color cut more promising since their simulations are not fine tuned to match the observed properties of dwarfs. Since we do not visually inspect all UDGs and UPGs in our sample, we take the color cut from \citet{CarlstenELVES2022} to define our quenched galaxies: galaxies that are redder than $g-i > -0.067 M_V - 0.23$ are defined as quenched, whereas those bluer galaxies are thought to be star-forming. However, one caveat is that this criterion was derived from LV satellites in \citet{CarlstenELVES2022} and the bulk of them have higher surface brightness and less extended than our UDGs and UPGs. It is not guaranteed that the same color cut would work in our regime. \footnote{\jiaxuan{Add some robustness tests, such as ``Changing the intercept of the color cut to -0.28 makes the overall quenched fraction higher, but does not change the trend of quenched fraction as a function of stellar mass.''}}

@@

describe the quenched fraction in percentage, e.g., half of them are quenched at 1e8.

@@

\subsection{High and constant quenched fraction of UDGs is an artifact}
In Section \ref{sec:quench}, we calculate the quenched fraction as a function of stellar mass for UDGs (left panels in Figure \ref{fig:qfrac}). We find that the UDG quenched fraction is reasonably a constant over $>1$ dex in stellar mass, and is higher than the quenched fractions of UPGs and satellite population of MW analogs. In this subsection, we argue that UDG's high and constant quenched fraction is merely an artifact due to its contrived definition.

@@

\subsection{UDG and UPG formation}
In this paper, one surprising finding is that although UPGs have extremely large size for their stellar mass, they have similar quiescent fraction as normal satellites in MW analogs. In ELVES, \citet{ELVES-I} find that the mass-size relation of satellites in MW analogs does not depend on the morphology (equivalently color) of satellites. They interpret this trend as that the quenching and morphological transformation involves very mild size evolution. These evidences might indicate that quenching and size growth must happen separately. Then the question is what physics is responsible for the two processes respectively. In this section, we first review the quenching of MW satellites and the mechanisms for puffing up normal dwarfs to produce UDGs (UPGs). Then we combine the results presented in Section \ref{sec:results} and discuss possible implications on the formation and quenching of UDGs and UPGs. 
\footnote{\jiaxuan{Greene et al. (in prep) also find no trend of quenched fraction on the size?}}

%%%%%%%%%%%%%%%%%%%
Isolated dwarf galaxies in the field are known to be star-forming \citep{Geha2012}, but satellites in the Local Group and Local Volume are more quiescent \citep{Grcevich2009,Spekkens2014,Wetzel2015,Putman2021,CarlstenELVES2022}. This difference hints that the MW environments play an important role on regulating the star formation activities in dwarf galaxies. There have been extensive works using simulations to study the quenching of satellites in MW analogs \citep[e.g.,][]{Simpson2018,Buck2019,Garrison-Kimmel2019,Simons2020,Joshi2021,Akins2021,Karunakaran2021,Font2022,Samuel2022}, and many of them are able to reproduce the observed quenched fraction in the Local Group and in the Local Volume. 
% the satellites in MW analogs are mostly quiescent below $M_\star\approx 10^7\ M_\odot$, and the quenched fraction drops as increasing stellar mass.

In theory, satellite galaxies in MW analogs are accreted from the field, therefore quenching can happen either before or after the infall to MW host. In the former case, isolated dwarf galaxies can be quenched by bursty stellar feedback \citep[e.g.,][]{Bahe2015,Kazantzidis2017,ElBadry2018}, reionization \citep[e.g.,][]{Bullock2000,Benson2002,Somerville2002,Tollerud2018,Applebaum2021}, or being backsplash satellites \citep{Simpson2018,Benavides2021}. If a satellite is star-forming before infall, the ram pressure \citep[e.g.,][]{GunnGott1972,McCarthy2008,Simpson2018,Tremmel2020,Samuel2022} experienced by the satellite as it move through the hot circumgalactic medium (CGM) of the host could remove the gas and cause the cessation of star formation. It has been shown that ram pressure stripping can 
efficiently quench galaxies at $M_\star < 10^{8}\ M_\odot$ \citep{Fillingham2016,Simpson2018,Buck2019}. Tidal stripping can also deprive the gas from the satellite or stop the gas accretion and starve the galaxy \citep{Simpson2018}, but strong tidal interactions are needed because dark matter needs to be deprived before significant gas loss happens. Another quenching mechanism is the so-called preprocessing in a lower-mass group prior to the infall into a MW analog \citep{Wetzel2015b,Jahn2022,Samuel2022}. Satellites might be quenched in the previous group via ram pressure striping or tidal interactions, and gas distribution can also be altered in preprocessing even without being quenched. \citet{Samuel2022} show that 40\% of satellites were preprocessed before the infall to MW. Some studies also show that the paired host environments might enhance the quenching of satellites (\citealt{Garrison-Kimmel2019,Putman2021}, but also see \citealt{Samuel2022}).

For UDGs and UPGs in MW analogs, they have extremely large size compared with normal satellites with their masses. Many mechanisms are proposed to explain the large size of UDGs (UPGs) in simulations. For isolated UDGs, bursty stellar feedback can cause the stellar component of dwarf galaxy to expand \citep[e.g.,][]{DiCintio2017,Chan2018,Martin2019,Jiang2019,Carleton2019}. Analytical models also support a scenario where UDGs originates from a population of dwarf galaxies residing in halos with higher spin \citep{Dalcanton1997,Amorisco2016,Rong2017,Liao2019}. \citet{Wright2021} show that field UDGs can also be formed from major mergers that cause star formation to
migrate outward. For UDGs in groups and clusters, environmental effects are more pronounced and tidal interactions are believed to puff up normal dwarfs. \citet{Jiang2019} find that the UDGs in groups are either accreted as UDGs from field or accreted as normal dwarfs but get tidally heated near the orbital pericenter. \citet{Tremmel2020} study UDG formation in cluster environment and find UDGs with higher mass ($M_\star > 10^{8}\ M_\odot$) are puffed up suddenly due to tidal heating at pericenter, but UDGs with lower mass ($M_\star \sim 10^{7.5}\ M_\odot$) are gradually puffed up due to adiabatic expansion as a response to the mass loss from tidal stripping and ram pressures stripping. \citet{Kado-Fong2021} study the intrinsic shape of UDGs in observations and find the inferred shapes agree with merger scenario for field UDGs \citep{Wright2021} and tidal heating/adiabatic expansion for group/cluster UDGs \citep{Jiang2019,Tremmel2020}.

For UDGs in groups, since they probably experienced strong tidal interactions, it is not obviously clear whether they are quenched mainly by tidal stripping or ram pressure stripping. 
% In the tidal stripping scenario, dark matter needs to be removed before a significant fraction of gas are deprived by tidal force, indicating a longer quenching timescale. 
\citet{Jiang2019} argue that if tidal stripping dominates the quenching, the satellite closer to the host will have smaller stellar mass and effective radius since stars are also stripped. They find no such trend in simulation, but find that the average color is redder when the UDGs closer to the host, supporting that ram pressure stripping is more dominant. In Section \ref{sec:size_distr}, we split the sample into two radial distance bins, and we do not find significant change in size or stellar mass distribution as UDGs get closer to the host.

In Auriga simulations, \citet{Simpson2018} show that the quenched fraction of satellites $M_\star < 10^8\ M_\odot$ have a strong dependence on the distance to the hosts. Using the FIRE-2 simulation, \citet{Samuel2022} also show that the quenched fraction increase as the distance to the host decreases, and hosts with higher CGM mass have more quenched satellites. In Figure \ref{fig:qfrac}, by comparing UDGs and UPGs at two radial bins, we find the very similar trends that UDGs (UPGs) closer to the host are more quiescent. In the ram pressure stripping picture, this is because the density of CGM increases when being closer to the host and the ram pressure is stronger. Together with the null trend on size and stellar mass distribution as UDGs (UPGs) get closer to their hosts, our findings support that ram pressure stripping is the major quenching channel for UDGs (UPGs) in MW analogs.

%Combining all these factors together, returning backsplash galaxies can also exhibit as other satellites.

If UDGs and UPGs are quenched by ram pressure stripping, how are they puffed up? UPGs can be puffed during the infall or prior to the infall. If UPGs are accreted into MW hosts from the field as normal dwarfs and puffed up due to tidal heating, we will expect a higher quenched fraction since they must undergo more violent tidal interactions, and the ram pressure is also stronger at distances closer to the host. If they are puffed up due to a longer term adiabatic expansion due to the mass loss, the expansion should also be proportional to the amount of gas loss, which translates to quenched fraction. Therefore, tidal puffing during infall seems to contradict with the normal quenched fraction as seen in Figure \ref{fig:qfrac}.

It is more feasible that UPGs are already puffed up before accretion and remain being star-forming. They are quenched by ram pressure stripping in the MW host with other normal satellites. In order to achieve similar quenched fraction as normal satellites, field UPGs prior to infall have to be star forming. Bursty star formation, early merger, higher halo spin, and preprocessing can all puff up the field dwarfs to become UPGs, and some of them seem not to quench the UPGs. \citet{Wright2021} show that early merger does not significant change the total SFR. \citet{Samuel2022} show that preprocessing can marginal boost the quenched fraction by 20\% at $M_\star \approx 10^8\ M_\odot$ and merely change the quenched fraction for lower mass satellites. Although it is not clear whether the high halo spin is the cause or the consequence of UPG formation, it is probable that higher halo spin does not quench a UPG. If any of these mechanisms can only increase the size to make UPGs but not quench them, it provides a plausible way to produce the observed trend in Fig. \ref{fig:qfrac} where quenching and size growth are decoupled.
However, the efficiency of ram pressure stripping also depends on the stellar mass density which provides the restoring force. The diffuse stellar component might make UPGs more vulnerable to ram pressure stripping and tidal stripping. It is also possible that UPGs in MW analogs are a mixture of field UPGs and normal dwarfs puffed up during infall. Using our observations, it is hard to sort out detailed mechanisms for quenching and size growth of UPGs. More detailed studies in observation and simulation are needed to solve this puzzle.

% \citet{Wright2021} observed a temporary boost in spin after the early merger. \citet{Liao2019} find that blue field UDGs indeed have higher halo spins. If field UDGs (UPGs) are born in high spin halos, they might have similar SFR as normal field dwarfs.

% \citet{Samuel2022} show that 40\% of satellites were in another group prior to the infall to MW, and preprocessed satellites have higher quenched fraction, and 10\% were quenched during preprocessing. 
\jiaxuan{Do we need to discuss splashback somewhere?}

% \jiaxuan{But UDGs have higher HI density?}

% splashback: they must be quenched in their prior infall.

% \citet{Samuel2022} find that HI gas can form a concentrated core after ram-pressure stripping and be there for quite long. \citet{Font2022} find LSB galaxies in FIRE-2 have concentrated stellar core. Oh, but the core might not be dense enough to form stars.

% Galaxy conformity: mass-size independent of color, jenny's quenched fraction independent of morph.

% Scenarios:
% 1. violent environmental effects: radial orbits, back-splash both explain large size and quenching. But according to Bernavidis et al, the quenched fraction drops with increasing mass.

% 2. accreting field UDGs, and quench them due to ram pressure stripping. But how long do accreted UDGs last before being destroyed? Need to look at ram pressure stripping literature.

% Homework:
% 1. Move the stellar mass bin definition, try to match with red/blue, spiral/elliptical plot. If cannot match, does this hint the formation scenarios?
% 2. How significant do we detect blue UDGs? 
% 3. Split into redshift bins

% UDG rejuvenate after ram pressure stripping?

% calculate probability of being a backsplash satellite (e.g., splashed to 1.5 Mpc)
% Compare spatial distribution with normal dwarf? iF they are indeed high-spin tail of normal dwarfs. 
% Host-to-host variation

From \citet{Wright2021}, We can take this into our paper: At high mass (M⋆ > 108 M⊙), most galaxies have reff > 1.5 kpc. UDGs are therefore those galaxies that are unusually LSB. At low mass (M⋆ < 107.5 M⊙), most galaxies have μ0,g > 24 mag arcsec−2. UDGs are therefore those galaxies that have unusually large effective radii. It is only really at intermediate masses that both the central surface brightness and the effective radius criteria are relevant.

@@

@@

\section{Summary}
\begin{itemize}
    \item we advocate studies for the quenched fraction as a function of size in simulations

    \item Future works: nucleation, intrinsic shape, LSBG cross-correlation, unsupervised machine learning, color gradient, non-par shapes. 
\end{itemize}
